A Plasma Torus Around a Young Low-Mass Star

### 2025-01-26 paper/ms.tex
update

\begin{table*}
\scriptsize
\setlength{\tabcolsep}{2pt}
\centering
\caption{Selected System Parameters for TIC~141146667}
\label{tab:params}
\begin{tabular}{llcc}
\hline \hline
Parameter & Description & Value & Source\\
\hline 
%
$T_{\rm eff}$\dotfill                   & Effective Temperature (K) \hspace{9pt}\dotfill                 & 2972 $\pm$ 40    & 1 \\
%
%$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
%
$R_\star$\dotfill                       & Stellar radius ($R_\odot$)\dotfill                             & 0.42$\pm$0.02    & 1 \\
%
$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & XXX  & 6 \\
%
% TODO: this value is a quicklook from calc_hires_systemic_velocities.py, rj537.166.fits.  should average over whole timeseries.
$\gamma$\dotfill                       & Systemic radial velocity (\kms)\dotfill                         & 2.2 $\pm$ 0.9  & 1  \\
%
Age                                     & Adopted stellar age (Myr)\dotfill                              & YYY  & 8 \\
%
Spec. Type\dotfill                      & Spectral Type\dotfill                                          & MXV              & X \\
%
$P_{\rm rot}$\dotfill                   & Rotation period (hr)\dotfill                                   & $3.930\pm 0.001$ & X \\
%
$v_{\rm eq}$\dotfill		                & Equatorial velocity \dotfill                                   &  130$\pm$4       & - \\
                                        & \hspace{3pt} ($2\pi R_\star/P_{\rm rot}$) (\kms)	             &                      \\
%
$v_{\rm eq}\sin{i_\star}$\dotfill		    & Projected rotational\dotfill                                   &  128$\pm$3       & X \\
                                        & \hspace{3pt} velocity (\kms)	                                 &                      \\
%
$i_\star$\dotfill                       & Stellar inclination (deg)\dotfill                              & 	XXX             & X \\
%
$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.6 \pm X.X$   & X \\
%
$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   &  XXX             & X \\
                                        & \hspace{3pt} radius ($R_\star$)	                               &                      \\
%
$a_1$\dotfill                           & Clump 1 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  2.1-2.7         & X \\
$a_2$\dotfill                           & Clump 2 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  3.9             & X \\
%
$\langle$EW$_{\rm H\alpha}$$\rangle$    & Time-averaged H$\alpha$ line core                              &  X.X             & Y \\ 
                                        & \hspace{3pt} equivalent width (\AA)	                           &                      \\
Range(EW$_{\rm H\alpha}$)$_1$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 1 (\AA)	                               &                      \\
Range(EW$_{\rm H\alpha}$)$_2$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 2 (\AA)	                               &                      \\
\hline
\end{tabular}
\begin{flushleft}
\footnotesize{ \textsc{NOTE}---
%$^\dagger$ The GAIADR2 and GAIAEDR3 identifiers for Kepler 1627A are identical.  The secondary
is not resolved in the Gaia point source catalog.
$^*$ Given only $v\sin i$ and $2\pi R_\star/P_{\rm rot}$, $\cos i=0.11^{+0.11}_{-0.08}$.
Provenances are:
1: SED fit\cite{Bouma2024},
%$^6$Cluster isochrone (MIST adopted; PARSEC compared for quoted
%  uncertainty),
2: TESS light curve,
%  3: HIRES spectra and specmatch(?),
%$^8$Pre-main-sequence CAMD interpolation (Section~\ref{sec:camd}),
}
\end{flushleft}
\vspace{-0.5cm}
\end{table*}

@@

{\it Radial Velocity}---We measured the radial velocities of
TIC~141146667 using a pipeline that we for rapidly rotating
stars, including M dwarfs.
%TODO: describe

Standard precision radial velocity pipelines
fail on stars like  due to their extreme line broadening.

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (km s$^{-1}$ ) & & \\
    $V$ (km s$^{-1}$ ) & & \\
    $W$ (km s$^{-1}$ ) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & todo & This work\\
    $i_\star$($^\circ$) & todo & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
%$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
%
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & todo & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}
\end{methods}

>>>>

\begin{table*}
\scriptsize
\setlength{\tabcolsep}{2pt}
\centering
\caption{Selected System Parameters for TIC~141146667}
\label{tab:params}
\begin{tabular}{llcc}
\hline \hline
Parameter & Description & Value & Source\\
\hline 
%
$T_{\rm eff}$\dotfill                   & Effective Temperature (K) \hspace{9pt}\dotfill                 & 2972 $\pm$ 40    & 1 \\
%
%$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
%
$R_\star$\dotfill                       & Stellar radius ($R_\odot$)\dotfill                             & 0.42$\pm$0.02    & 1 \\
%
$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & XXX  & 6 \\
%
$\gamma$\dotfill                        & Systemic radial velocity (\kms)\dotfill                        & 0.61 $\pm$ 1.47  & 1  \\
%
Age                                     & Adopted stellar age (Myr)\dotfill                              & YYY  & 8 \\
%
Spec. Type\dotfill                      & Spectral Type\dotfill                                          & MXV              & X \\
%
$P_{\rm rot}$\dotfill                   & Rotation period (hr)\dotfill                                   & $3.930\pm 0.001$ & X \\
%
$v_{\rm eq}$\dotfill		                & Equatorial velocity \dotfill                                   &  130$\pm$4       & - \\
                                        & \hspace{3pt} ($2\pi R_\star/P_{\rm rot}$) (\kms)	             &                      \\
%
$v_{\rm eq}\sin{i_\star}$\dotfill		    & Projected rotational\dotfill                                   &  128$\pm$3       & X \\
                                        & \hspace{3pt} velocity (\kms)	                                 &                      \\
%
$i_\star$\dotfill                       & Stellar inclination (deg)\dotfill                              & 	XXX             & X \\
%
$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.6 \pm X.X$   & X \\
%
$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   &  XXX             & X \\
                                        & \hspace{3pt} radius ($R_\star$)	                               &                      \\
%
$a_1$\dotfill                           & Clump 1 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  2.1-2.7         & X \\
$a_2$\dotfill                           & Clump 2 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  3.9             & X \\
%
$\langle$EW$_{\rm H\alpha}$$\rangle$    & Time-averaged H$\alpha$ line core                              &  X.X             & Y \\ 
                                        & \hspace{3pt} equivalent width (\AA)	                           &                      \\
Range(EW$_{\rm H\alpha}$)$_1$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 1 (\AA)	                               &                      \\
Range(EW$_{\rm H\alpha}$)$_2$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 2 (\AA)	                               &                      \\
\hline
\end{tabular}
\begin{flushleft}
\footnotesize{ \textsc{NOTE}---
%$^\dagger$ The GAIADR2 and GAIAEDR3 identifiers for Kepler 1627A are identical.  The secondary
is not resolved in the Gaia point source catalog.
$^*$ Given only $v\sin i$ and $2\pi R_\star/P_{\rm rot}$, $\cos i=0.11^{+0.11}_{-0.08}$.
Provenances are:
1: SED fit\cite{Bouma2024},
%$^6$Cluster isochrone (MIST adopted; PARSEC compared for quoted
%  uncertainty),
2: TESS light curve,
%  3: HIRES spectra and specmatch(?),
%$^8$Pre-main-sequence CAMD interpolation (Section~\ref{sec:camd}),
}
\end{flushleft}
\vspace{-0.5cm}
\end{table*}

@@

{\it Radial Velocity}---We measured radial velocities of
TIC~141146667 from each of our spectra using a pipeline that we
developed specifically for rapidly rotating stars.
Our method is based on template-matching against synthetic spectra
produced by the PHOENIX stellar atmosphere code \cite{Husser2013}.
We used the PHOENIX models with solar metallicity and alpha element
abundances, and calibrated our absolute velocity zero-points as well
as viable orders for velocity measurement using the
standard stars described by \cite{Chubak2012}.
We used velocity standards spanning spectral types from G2 to M4 (Barnard's
Star), irrespective of rotation rate.
We used \texttt{barycorrpy} \cite{Kanodia2018} to calculate the
velocity corrections due to Earth's motion around the solar system
barycenter and due to Earth's daily rotation about its axis.
Our analysis pipeline reproduces the systemic velocities reported by
\cite{Chubak2012} for their velocity standards with an RMS of
0.66\,\kms.

For TIC~141146667, we measured the velocities from each of our HIRES
spectra using regions near the K~I (7700\,\AA) resonance line and
three TiO bandheads (5160\,\AA, 5450\,\AA, and 5600\,\AA).  These
regions were selected because they provided
the best matches between the synthetic and observed spectra.  The
resulting
redshift measurements were averaged together.
The star's 130$\pm$4\,\kms\ equatorial velocity and near-edge on
viewing orientation yielded a line broadening $\Delta
\lambda$$\approx$3\,\AA.
We used the scatter of resulting velocity measurements between 
orders to assign the RV uncertainty at each epoch.
The uncertainty-weighted mean systemic velocity over all epochs on UT
2024-02-17 was
$\gamma$=0.6$\pm$1.5\,\kms.
The relative radial velocities about this mean are given in
Table~\ref{tab:rv}.
% from calc_max_amplitude_sinusoid_rv_timeseries.py
This radial velocity time-series rules out any significant periodicity
in the radial velocities at the rotation period exceeding
semi-amplitudes of 2.85\,\kms\ (at 3$\sigma$ confidence).
% from K_to_msini.py, assuming Mstar = 0.18
This in turn sets an upper limit on the mass of any putative
companions at this period of $m \sin i $$<$2.4\,$M_{\rm Jup}$.

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (\kms) & & \\
    $V$ (\kms) & & \\
    $W$ (\kms) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & todo & This work\\
    $i_\star$($^\circ$) & todo & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
    %$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & todo & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{lcr}
  %\setlength{\tabcolsep}{2pt}
  \hline 
  \hline 
  Time [JD$_{\rm UTC}$] & RV (\kms) & $\sigma_{\rm RV}$ (\kms) \\
  % from tables/tab_TIC141146667_rel_rv.tex
  \hline 
  60357.450329 & 1.72 & 5.86 \\
  60357.461255 & -5.41 & 2.37 \\
  60357.472181 & -1.21 & 2.64 \\
  60357.483109 & 2.83 & 2.87 \\
  60357.494030 & 6.52 & 7.53 \\
  60357.504949 & -3.0 & 1.44 \\
  60357.515873 & 0.01 & 1.21 \\
  60357.526794 & -0.38 & 7.03 \\
  60357.537717 & -3.92 & 2.71 \\
  60357.548639 & 7.92 & 6.75 \\
  60357.559566 & 4.94 & 8.84 \\
  60357.570487 & -3.26 & 3.06 \\
  60357.581408 & 0.83 & 1.34 \\
  60357.592330 & 1.4 & 8.24 \\
  60357.603251 & -8.05 & 3.94 \\
  60357.614172 & -3.26 & 3.07 \\
  60357.625095 & -3.84 & 7.55 \\
  60357.636019 & -1.6 & 2.26 \\
  60357.646940 & 0.82 & 2.91 \\
  60357.657861 & 3.53 & 3.95 \\
  60357.668781 & 5.2 & 12.14 \\
  \hline
  \end{tabular}
  \caption{TIC~141146667 radial velocities.}
  \label{tab:rv}
\end{table}

\end{methods}
